Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms

### 2018-07-25 paper/paper.tex
Betterer

\usepackage{showframe}

@@

@@

\definecolor{highlight}{rgb}{0.819608, 0.733333, 0.34902}

@@

The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph inside a
larger ``target'' graph. Although \NP-complete, by combining design techniques from artificial
intelligence with careful algorithm engineering, modern subgraph isomorphism solvers can often
produce exact solutions quickly even on graphs with thousands of vertices. The current single
strongest subgraph isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to
direct a constraint programming style search
\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
heuristic is much better than branching randomly, but is still far from perfect.  We could try to
introduce tiebreaking mechanisms and similar refinements to strengthen the heuristic. However, given
the extreme cost of making an incorrect branching choice early in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which performs poorly), and a new form of search
(which is extremely beneficial) based around restarts and nogood recording. To explore different
parts of the search space when restarting, we introduce a novel value-ordering heuristic, which adds
a small amount of randomness to the ``highest degree first'' heuristic.  The effects of these
changes are shown in the final plot of \cref{figure:scatter-random}: we can make a state-of-the-art algorithm
perform much better on a large number of satisfiable instances, whilst performing worse only rarely
on satisfiable instances, and never on unsatisfiable instances.  Further, these properties are
\emph{not} shared by discrepancy searches (\cref{figure:scatter-dds}) or by random
value-ordering with restarts (the first two plots of \cref{figure:scatter-random}). Finally, we
show that this strategy is also effective in an optimisation setting, producing benefits in two
maximum common induced subgraph algorithms.

@@

Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself. The
\emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. When the graph is clear from the context, we omit
the $G$ subscripts. As per \citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
simple paths of length exactly $\ell$ between vertices $v$ and $w$ in $G$. Finally, we define a
function $\FuncSty{compatible}(v, w)$ to mean that mapping pattern vertex $v$ to target vertex $w$
is feasible based upon both loop constraints and the neighbourhood degree sequence relation of
\citet{DBLP:journals/constraints/ZampelliDS10}.

The non-induced subgraph isomorphism problem is to find an injective mapping from the vertices of a
pattern graph $\mathcal{P}$ to a target graph $\mathcal{T}$, such that adjacent vertices in
$\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$. We write $v \mapsto v'$ to mean that
pattern vertex $v$ is mapped to target $v'$ under such a mapping.

@@

\begin{description}
    \item[BVG(r), M4D(r), and Rand] are randomly generated graphs using different models (bounded
        degree, regular mesh, and uniform), where each pattern is a permuted random connected subgraph
        of the target (and so each instance is satisfiable) \cite{DBLP:journals/pami/CordellaFSV04}.
        These benchmark instances are widely used, but have unusual properties and so broad
        conclusions should not be drawn based solely upon behaviour on these instances
        \cite{DBLP:journals/jair/McCreeshPST18}.
    \item[SF] contains randomly generated scale-free graphs using a similar method
        \cite{DBLP:journals/constraints/ZampelliDS10}.
   \item[LV] consists of various kinds of graph gathered by Larrosa and Valiente
       \cite{DBLP:journals/mscs/LarrosaV02} from the Stanford Graph Database. We include both the
        50 small graphs, and the 50 larger graphs.
    \item[Phase] contains hard crafted instances that lie near the satisfiable / unsatisfiable phase
        transition \cite{DBLP:journals/jair/McCreeshPST18}.
    \item[PR] contains graphs generated from segmented images, corresponding to a computer vision
        problem \cite{DBLP:journals/pr/SolnonDHJ15}.
    \item[Images and Meshes] contain graphs representing 2D segmented images and 3D object models
        respectively, again representing a computer vision problem
        \cite{DBLP:journals/cviu/DamiandSHJS11}.
\end{description}

@@

\item[Biased] We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \]\end{description}

The biased heuristic continues to prefer vertices of high degree, but will give the same chance of
being selected next to two vertices of equal degree.  It also introduces an element of randomness,
which is needed to make restarts have an effect.  We note that it resembles the well-known
\emph{softmax} weighting scheme applied to the degree heuristic, although usually \emph{softmax}
uses a base of $e$ rather than $2$.  Experiments showed that in this context, the choice of base is
not important to the search space size; however, a base of $2$ avoids the need to use floating point
arithmetic, which was a substantial overhead.

@@

The effectiveness of this approach raises a number of interesting questions and avenues for possible
future research. For example, will this same technique work for other problems where we have a good
value-ordering heuristic? Are there further benefits to be had from also modifying variable-ordering
heuristics? Are there theoretical rather than just empirical reasons for using the
\emph{softmax}-like bias scheme? And how can such a search process be parallelised?

>>>>

% \usepackage{showframe}

@@

% cref style
\crefname{figure}{Figure}{Figures}
\Crefname{figure}{Figure}{Figures}
\crefname{section}{Section}{Sections}
\Crefname{section}{Section}{Sections}
\crefname{proposition}{Proposition}{Propositions}
\Crefname{proposition}{Proposition}{Propositions}
\crefname{corollary}{Corollary}{Corollaries}
\Crefname{corollary}{Corollary}{Corollaries}

\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}

@@

\definecolor{highlight}{rgb}{0.529412, 0.74902, 0.466667}

@@

The subgraph isomorphism problem is to decide whether a copy of a small ``pattern'' graph occurs
inside a larger ``target'' graph. Although \NP-complete, by combining design techniques from
artificial intelligence with careful algorithm engineering, modern subgraph isomorphism solvers can
often produce exact solutions quickly even on graphs with thousands of vertices. The current single
strongest subgraph isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to
direct a constraint programming style search
\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
heuristic is much better than branching randomly, but is still far from perfect.  We could try to
introduce tiebreaking mechanisms and similar refinements to strengthen the heuristic. However, given
the extreme cost of making an incorrect branching choice early in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which performs poorly), and a new form of search
(which is extremely beneficial) based around restarts and nogood recording. To explore different
parts of the search space when restarting, we introduce a novel way of adapting a value-ordering
heuristic to add a very small amount of randomness. The effects of these changes are shown in the
final plot of \cref{figure:scatter-random}: we can make a state-of-the-art algorithm perform much
better on a large number of satisfiable instances, whilst performing worse only rarely on
satisfiable instances, and never on unsatisfiable instances.  Further, these properties are
\emph{not} shared by discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering
with restarts (the first two plots of \cref{figure:scatter-random}). Finally, we show that this
strategy is also effective in an optimisation setting, producing benefits in two maximum common
induced subgraph algorithms.

@@

Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself. The
\emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood. When the graph is clear from the context, we omit
the $G$ subscripts. As per \citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with
vertex set $\vertexset(G)$, and with edges between vertices $v$ and $w$ if there are at least $n$
simple paths of length $\ell$ between vertices $v$ and $w$ in $G$.

The non-induced subgraph isomorphism problem is to find an injective mapping from the vertices of a
pattern graph $\mathcal{P}$ to a target graph $\mathcal{T}$, such that adjacent vertices in
$\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$ (including that vertices with loops
in $\mathcal{P}$ may only be mapped to vertices with loops in $\mathcal{T}$). We write $v \mapsto
v'$ to mean that pattern vertex $v$ is mapped to target $v'$ under such a mapping. Finally, we
define a function $\FuncSty{compatible}(v, w)$ to mean that mapping $v \mapsto w$ is feasible based
upon both the loop rule, and the neighbourhood degree sequence relation of
\citet{DBLP:journals/constraints/ZampelliDS10}.

@@

\begin{description}
    \item[BVG(r), M4D(r), and Rand] are families of randomly generated graphs using different models (bounded
        degree, regular mesh, and uniform), where each pattern is a permuted random connected subgraph
        of the target (and so each instance is satisfiable) \cite{DBLP:journals/pami/CordellaFSV04}.
        These benchmark instances are widely used, but have unusual properties and so broad
        conclusions should not be drawn based solely upon behaviour on these instances
        \cite{DBLP:journals/jair/McCreeshPST18}.
    \item[SF] contains randomly generated scale-free graphs using a similar method
        \cite{DBLP:journals/constraints/ZampelliDS10}.
   \item[LV] consists of various kinds of graph gathered by Larrosa and Valiente
       \cite{DBLP:journals/mscs/LarrosaV02} from the Stanford Graph Database. We include both the
        50 small graphs, and the 50 larger graphs.
    \item[Phase] contains hard crafted instances that lie near the satisfiable / unsatisfiable phase
        transition \cite{DBLP:journals/jair/McCreeshPST18}.
    \item[PR] contains graphs generated from segmented images, corresponding to a computer vision
        problem \cite{DBLP:journals/pr/SolnonDHJ15}.
    \item[Images and Meshes] contain graphs representing 2D segmented images and 3D object models
        respectively, again representing a computer vision problem
        \cite{DBLP:journals/cviu/DamiandSHJS11}.
\end{description}

@@

\item[Biased] We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This is a new
heuristic, which continues to prefer vertices of high degree, but will give the same chance of
being selected next to two vertices of equal degree.  It also introduces an element of randomness,
which is needed to make restarts have an effect. \end{description}

@@

The effectiveness of this approach raises a number of interesting questions and avenues for possible
future research. For example, will this same technique work for other problems where we have a good
value-ordering heuristic? Are there further benefits to be had from also modifying variable-ordering
heuristics? Are there theoretical rather than just empirical reasons for using an exponential bias
scheme? And how can such a search process be parallelised?
